
---
title: "regression_analysis"
author: "Philip Ashton"
date: "`r Sys.Date()`"
output: html_document
---

# The purpose of this script is to
# 1. make a nice dataframe that looks like this
# date, number of introductions, restriction level, country
# do a regression analysis, including autocorrelation

# import the packages required for this script
```{r}
library(readr)
library(lubridate)
library(dplyr)
library(stringr)
library(nlme)
```

# Sources
```{r sources}
source("/Users/flashton/Dropbox/COVID/AAP_introductions/scripts/utils.R")

```
# calc the intros per day

Did this again, as the other way was quite VOC orientated (only considered cases of specific VOCs), this will count any introduction of any voc.

```{r basics}

countries_to_include <- c("MWI", "KEN", "ZAF", "VNM", "THA", "IDN", "KHM")
national_cases_handle <- "/Users/flashton/Dropbox/COVID/our_world_in_data/2022.12.05/owid-covid-data.csv"
restrictions_handle <- "/Users/flashton/Dropbox/COVID/our_world_in_data/2022.12.05/international-travel-covid.csv"

national_cases <- read_in_national_case_numbers(national_cases_handle, countries_to_include)
#View(national_cases)
restrictions <- read_in_restrictions(restrictions_handle, countries_to_include)
restrictions <- restrictions %>% rename(iso_code = Code, date = Day)

get_intros_per_day <- function(per_sample_intro_handle, country_iso){
  per_sample_intros <- read_delim(per_sample_intro_handle, delim = "\t", escape_double = FALSE, trim_ws = TRUE)
  per_sample_intros <- per_sample_intros %>% mutate(date = lubridate::parse_date_time(str_split_i(sample, "\\|", -1), c('ymd', 'ym', 'y')))

  # View(per_sample_intros)
  first_introduction <- per_sample_intros %>% arrange(date) %>% distinct(introduction_node, .keep_all = TRUE)
  #print(date_range[1])
  #print(date_range[2])
  first_introduction$date <- as.Date(first_introduction$date)
  intros_per_day <- first_introduction %>% group_by(date) %>% summarise(number_introductions = n()) %>% mutate(country = country_iso)
  
  return(intros_per_day)
  #write_csv(intros_per_day, "/Users/flashton/Dropbox/COVID/AAP_introductions/results/2023.06.06/malawi.intros_per_day.csv")
}

malawi_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/malawian_introductions.excl_mwi_dups.txt", country_iso = "MWI")
cambodia_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/cambodian_introductions.txt", country_iso = "KHM")
indonesia_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/indonesian_introductions.txt", country_iso = "IDN")
kenya_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/kenyan_introductions.txt", country_iso = "KEN")
# laos_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/laotian_introductions.txt", country_iso = "LAO")
south_africa_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/southafrican_introductions.txt", country_iso = "ZAF")
thailand_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/thai_introductions.txt", country_iso = "THA")
vietnam_ipd <- get_intros_per_day("/Users/flashton/Dropbox/COVID/AAP_introductions/results/2022.12.02/vietnamese_introductions.txt", country_iso = "VNM")
```

# Combine the introductions per day with the restirction and case number data
```{r combine_ipd_restrictions_case_numbers}
# rbind malawi and cambodia
all_ipd <- rbind(malawi_ipd, cambodia_ipd, indonesia_ipd, kenya_ipd, south_africa_ipd, thailand_ipd, vietnam_ipd)
all_ipd <- all_ipd %>% rename(iso_code = country)
# View(all_ipd)

# malawi_ipd <- malawi_ipd %>% mutate(date = sampling_date)
cases_and_ipd <- national_cases %>% left_join(all_ipd, by = c("iso_code", "date")) %>% mutate(number_introductions = ifelse(is.na(number_introductions), 0, number_introductions))
cases_ipd_restrictions <- cases_and_ipd %>% left_join(restrictions, by = c("iso_code", "date")) %>% select(iso_code, continent, location, date, new_cases, number_introductions, international_travel_controls)
cases_ipd_restrictions <- cases_ipd_restrictions %>% mutate(number_introductions_lead14 = lead(number_introductions, n = 14)) %>% mutate(number_introductions_lead7 = lead(number_introductions, n = 7))
write_csv(cases_ipd_restrictions, "/Users/flashton/Dropbox/COVID/AAP_introductions/scripts/cases_ipd_restrictions.csv")


```


# use lme to model the number of introductions per day, with autocorrelation, including date and international travel controls as covars

setup

```{r setup}
cases_ipd_restrictions <- read_csv("/Users/flashton/Dropbox/COVID/AAP_introductions/scripts/cases_ipd_restrictions.csv")
cases_ipd_restrictions$international_travel_controls <- as.factor(cases_ipd_restrictions$international_travel_controls)
cases_ipd_restrictions <- cases_ipd_restrictions %>% filter(!is.na(international_travel_controls))

```


## Malawi - whole pandemic

```{r malawi_whole_pandemic}

levels(cases_ipd_restrictions$international_travel_controls)

mwi_cases_ipd_restrictions <- cases_ipd_restrictions %>% filter(iso_code == "MWI")
# lme(number_introductions ~ date + international_travel_controls, random = ~ date, correlation = corAR1(form = ~ date), data = mwi_cases_ipd_restrictions)
m1 <- lme(number_introductions ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions)
summary(m1)
```
```{r malawi_whole_pandemic_lead14}
m1_lead14 <- lme(number_introductions_lead14 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions)
summary(m1_lead14)
```

```{r malawi_whole_pandemic_lead7}
m1_lead7 <- lme(number_introductions_lead7 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions)
summary(m1_lead7)
```

## Malawi - beta wave

No lag
```{r malawi_beta_wave}
mwi_cases_ipd_restrictions_beta <- mwi_cases_ipd_restrictions %>% filter(date > "2020-12-01" & date < "2021-04-01") 

m1_beta <- lme(number_introductions ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_beta)
summary(m1_beta)
```

Lead of 14 days on introductions
```{r malawi_beta_wave_lead14}
m1_beta_lead14 <- lme(number_introductions_lead14 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_beta)
summary(m1_beta_lead14)
```


Lead of 7 days on introductions
```{r malawi_beta_lead7}

m1_beta_lead7 <- lme(number_introductions_lead7 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_beta)
summary(m1_beta_lead7)
```

## Malawi - delta wave

No lag

```{r malawi_delta_wave}
mwi_cases_ipd_restrictions_delta <- mwi_cases_ipd_restrictions %>% filter(date > "2021-04-01" & date < "2021-10-01") 

m1_delta <- lme(number_introductions ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_delta)
summary(m1_delta)
```

Lead of 14 days on introductions
```{r malawi_delta_lead14}

m1_delta_lead14 <- lme(number_introductions_lead14 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_delta)
summary(m1_delta_lead14)
```


Lead of 7 days on introductions
```{r malawi_delta_lead7}

m1_delta_lead7 <- lme(number_introductions_lead7 ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = mwi_cases_ipd_restrictions_delta)
summary(m1_delta_lead7)
```

## all countries entire pandemic
```{r all_countries_entire_pandemic}
m2 <- lme(number_introductions ~ date + international_travel_controls, random = ~ date | iso_code, correlation = corAR1(form = ~ date | iso_code), data = cases_ipd_restrictions)
summary(m2)
```